---
title: "Fee Schedule Calculations"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = FALSE,
  echo       = TRUE, 
  message    = FALSE, 
  warning    = FALSE,
  error      = TRUE,
  comment    = "#>",
  dpi        = 600, 
  out.width  = "100%",
  fig.height = 14,
  fig.width  = 10
)
options(scipen = 999)
```


```{r pkgs, message=FALSE, warning=FALSE}
library(northstar)
library(dplyr)
library(tinytable)
```

# Physician Fee Schedule Amount Calculation

```{r echo=FALSE}
dplyr::tribble(
  ~Variables,  ~Descriptions,
  "$RVU_w$",   "Relative Value Unit (Physician Work)",
  "$GPCI_w$",  "Geographic Practice Cost Index (Physician Work)",
  "$RVU_p$",   "Relative Value Unit (Practice Expense)",
  "$GPCI_p$",  "Geographic Practice Cost Index (Practice Expense)",
  "$RVU_m$",   "Relative Value Unit (Malpractice)",
  "$GPCI_m$",  "Geographic Practice Cost Index (Malpractice)",
  "$CF$",      "Conversion Factor") |> 
  tt(caption = "Physician Fee Schedule Variables") |> 
  format_tt(markdown = TRUE) |> 
  style_tt(bootstrap_class = "table table-hover")
```

<br>

```{r echo=FALSE}
dplyr::tribble(
  ~Formulas,                                         ~Results,
  "$RVU_w(GPCI_w) + RVU_p(GPCI_p) + RVU_m(GPCI_m)$", "Total RVUs",
  "$\\text{Total RVUs} * CF$",                       "Participating Fee",
  "$\\text{Participating Fee} * 0.95$",              "Non-Participating Fee",
  "$(\\text{Participating Fee} * 1.15) * 0.95$",     "Limiting Charge") |> 
  tt(caption = "Physician Fee Schedule Formulas") |> 
  format_tt(markdown = TRUE) |> 
  style_tt(bootstrap_class = "table table-hover")
```

<br>

As there is a corresponding GPCI value for each of the three components of an RVU, the formula is essentially the [dot product](https://www.toppr.com/guides/maths-formulas/dot-product-formula/) of the RVU and GPCI vectors, multiplied by the Conversion Factor:

$ RVU \cdot GPCI = RVU_w(GPCI_w) + RVU_p(GPCI_p) + RVU_m(GPCI_m)$

```{r}
wrvu  = 6.26   # Work RVU
wgpci = 1      # Work GPCI

pgpci = 0.883  # Practice GPCI
prvu  = 4.36   # Practice RVU

mrvu  = 0.99   # Malpractice RVU
mgpci = 1.125  # Malpractice GPCI

cf    = 32.744 # Conversion Factor

rvus <- (wrvu * wgpci) + (prvu * pgpci) + (mrvu * mgpci)
rvus # Total RVUs

rvus * cf # Participating Fee
```

<br>

Using base r's matrix multiplication infix operator

```{r}
as.vector(c(wrvu, prvu, mrvu) %*% c(wgpci, pgpci, mgpci))
```

<br>

Using the `dot()` function from the `{pracma}` package

```{r}
pracma::dot(
  c(wrvu, prvu, mrvu), 
  c(wgpci, pgpci, mgpci))
```

<br>



```{r, results='asis'}
calculate_amounts(
  wrvu  = 6.26,
  nprvu = 7.92,
  fprvu = 4.36,
  mrvu  = 0.99,
  cf    = 32.744,
  wgpci = 1,
  pgpci = 0.883,
  mgpci = 1.125
)
```
