---
title: "Fee Schedule Calculations"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = FALSE,
  echo      = TRUE, 
  message   = FALSE, 
  warning   = FALSE,
  error     = TRUE,
  comment   = "#>",
  dpi       = 600, 
  out.width = "100%",
  fig.height = 14,
  fig.width = 10
)
options(scipen = 999)
```


```{r pkgs, message=FALSE, warning=FALSE}
library(northstar)
library(dplyr)
library(gt)
```

# Physician Fee Schedule Amount Calculation

|          |                                                   |
|----------|---------------------------------------------------|
| $RVU_w$  | *Work* Relative Value Unit                        |
| $GPCI_w$ | *Work* Geographic Practice Cost Index             |
| $RVU_p$  | *Practice Expense* Relative Value Unit            |
| $GPCI_p$ | *Practice Expense* Geographic Practice Cost Index |
| $RVU_m$  | *Malpractice* Relative Value Unit                 |
| $GPCI_m$ | *Malpractice* Geographic Practice Cost Index      |
| $CF$     | Conversion Factor                                 |

<br>

The __Participating amount__ is calculated as follows:

<br><br>

$$x = [(RVU_w \times GPCI_w) + (RVU_p \times GPCI_p) + (RVU_m \times GPCI_m)] \times CF$$

<br>

The __Non-participating amount__ is calculated as follows:

<br>

$$ y = 0.95x$$

<br>

The __Limiting charge__ is calculated as follows:

<br>

$$ z = 0.95(1.15x)$$

<br>

```{r}
calculate_amounts(
  wrvu  = 6.26,
  nprvu = 7.92,
  fprvu = 4.36,
  mrvu  = 0.99,
  cf    = 32.744,
  wgpci = 1,
  pgpci = 0.883,
  mgpci = 1.125
)
```
